---
title: "Making a 5 microstep GIF"
author: Sam Tyner
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```


## Where do we start? 

We start with some data that we want to model using `RSiena`. So, let's find some data and fit a simple model to it. 

```{r initializemodel}
library(RSiena)
friend.data.w1 <- as.matrix(read.table("../data/s50-network1.dat"))
friend.data.w2 <- as.matrix(read.table("../data/s50-network2.dat"))
friend.data.w3 <- as.matrix(read.table("../data/s50-network3.dat"))
drink <- as.matrix(read.table("../data/s50-alcohol.dat"))
friendship <- sienaDependent(
  array(c(friend.data.w1, friend.data.w2,
          friend.data.w3),
        dim = c(50, 50, 3)))
alcohol <- varCovar(drink)
mydata <- sienaDataCreate(friendship, alcohol)
myeffnull <- getEffects(mydata)
myalgorithm <- sienaAlgorithmCreate(projname = 's50_3')
set.seed(823746)
ansnull <- siena07(myalgorithm, data = mydata, effects = myeffnull,
                   returnChains = TRUE, returnDataFrame = TRUE,
                   returnDeps = TRUE, silent = TRUE, verbose = FALSE,
                   batch = TRUE)
```

## What's next?

Now that we have our fitted model, we want to look "underneath the hood" to see the microsteps that are going on underneath. Let's extract those microsteps. 

```{r getms}
library(netvizinf)
library(dplyr)
library(readr)
library(sna)
library(network)
ansnullchains <- get_chain_info(ansnull)
ansnullchains %>% 
  filter(period == 1) %>%  #only look at chains from wave 1 to wave 2
  group_by(rep) %>%
  select(rep, from = X4, to = X5) %>% 
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1)     # match the data vars 
  ) -> ansnullchainsw1w2

# create microstep data from 1 of the 1000 reps in the chain. 
library(geomnet)
wave1friends <- fortify(as.adjmat(friend.data.w1))
ms1 <- listMicrosteps(dat = wave1friends, 
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
# we're only going to do the first 5 microsteps. 
microsteps <- ms1[1:5]
```

## Get data in a form to visualize

Now, we want to create datasets that are appropriate for visualization using some functions from `netvizinf` and `tweenr`. 

```{r tweeen}
to_tween <- pretween(microsteps = microsteps, layoutparams = list(n = 50, directed = TRUE))
library(tweenr)
tweendat <- tween_states(data = to_tween, 18 ,1,"linear", nframes = 100)
# create some visualization variables. 
tweendat$addedge <- as.logical(tweendat$addedge)
tweendat$rmvedge <- as.logical(tweendat$rmvedge)
# color new edges
library(ggplot2)
library(animation)
library(gganimate)
p <- ggplot(data = unique(tweendat[, c("from", "x.from", "y.from",
                                  "addedge", "rmvedge", ".frame")]), 
                          aes(x=x.from, y = y.from, frame = .frame)) +
  geom_segment(data = tweendat, aes(x = x.from, y = y.from, xend = x.to, yend = y.to, frame = .frame, color = addedge)) + 
  geom_point(size=5, color = 'grey40') + 
  geom_text(aes(label=from), size = 3, alpha = .5, color = 'white') + 
  scale_colour_manual(values = c("black", "red")) + 
  scale_alpha_identity() + 
  theme_bw() 
animation::ani.options(interval = 1/2)
gganimate(p, "fivemicrosteps_attempt1.gif", title_frame = F)
```
