---
title: "Making a 5 microstep GIF"
author: Sam Tyner
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```


## Where do we start? 

We start with some data that we want to model using `RSiena`. So, let's find some data and fit a simple model to it. 

```{r initializemodel}
library(RSiena)
friend.data.w1 <- as.matrix(read.table("../data/s50-network1.dat"))
friend.data.w2 <- as.matrix(read.table("../data/s50-network2.dat"))
friend.data.w3 <- as.matrix(read.table("../data/s50-network3.dat"))
drink <- as.matrix(read.table("../data/s50-alcohol.dat"))
friendship <- sienaDependent(
  array(c(friend.data.w1, friend.data.w2,
          friend.data.w3),
        dim = c(50, 50, 3)))
alcohol <- varCovar(drink)
mydata <- sienaDataCreate(friendship, alcohol)
myeffnull <- getEffects(mydata)
myalgorithm <- sienaAlgorithmCreate(projname = 's50_3')
set.seed(823746)
ansnull <- siena07(myalgorithm, data = mydata, effects = myeffnull,
                   returnChains = TRUE, returnDataFrame = TRUE,
                   returnDeps = TRUE, silent = TRUE, verbose = FALSE,
                   batch = TRUE)
```

## What's next?

Now that we have our fitted model, we want to look "underneath the hood" to see the microsteps that are going on underneath. Let's extract those microsteps. 

```{r getms}
library(netvizinf)
library(dplyr)
library(readr)
library(sna)
library(network)
ansnullchains <- get_chain_info(ansnull)
ansnullchains %>% 
  filter(period == 1) %>%  #only look at chains from wave 1 to wave 2
  group_by(rep) %>%
  select(rep, from = X4, to = X5) %>% 
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1)     # match the data vars 
  ) -> ansnullchainsw1w2

# create microstep data from 1 of the 1000 reps in the chain. 
library(geomnet)
wave1friends <- fortify(as.adjmat(friend.data.w1))
ms1 <- listMicrosteps(dat = wave1friends, 
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
# we're only going to do the first 5 microsteps. 
microsteps <- ms1[1:5]
```

## Get data in a form to visualize

Now, we want to create datasets that are appropriate for visualization using some functions from `netvizinf` and `tweenr`. 

```{r tweeen}
to_tween <- pretween_vertices(microsteps = microsteps, layoutparams = list(n = 50))
library(tweenr)
tweendat <- tween_states(data = to_tween, 20 ,1,"linear", nframes = 100)
# create some visualization variables.
pte <- pretween_edges(microsteps)
# color new edges
library(ggplot2)
library(animation)
library(gganimate)
p <- ggplot(data = unique(tweendat[, c("from", "x.from", "y.from",
                                  "addedge", "rmvedge", ".frame")]), 
                          aes(x=x.from, y = y.from, frame = .frame)) +
  geom_segment(data = tweendat, aes(x = x.from, y = y.from, xend = x.to, yend = y.to, frame = .frame, color = addedge)) + 
  geom_point(size=5, color = 'grey40') + 
  #geom_text(aes(label=from), size = 3, alpha = .5, color = 'white') + 
  scale_colour_manual(values = c("black", "red"), guide = guides(color = "none")) + 
  scale_alpha_identity() + 
  theme_void() 
animation::ani.options(interval = 1/4)
gganimate(p, "fivemicrosteps_attempt1.gif", title_frame = F)
```

## My incredibly screwed up attempt to do this more efficiently 
```{r justabunchofjunk}
# baby edge disappear example

layout1 <- data.frame(x = c(0,1), y = c(0,1), id = c("V1", "V2"), ms = 0, rmvedge = c(T,F))
layout2 <- data.frame(x = c(0,1), y = c(1,0), id = c("V1", "V2"), ms = 1, rmvedge = F)

babytweend <- tween_states(list(layout1, layout2), 5,1,"linear", 5)

babyedge1 <- data.frame(from = "V1", to = "V2", rmvedge = TRUE, ms = 0)
babyedge2 <- data.frame(from = c("V1", "V2"), to = NA, rmvedge = FALSE, ms = 1)
babyedges <- rbind(babyedge1, babyedge2)

babystep1 <- merge(babyedges, babytweend, by.x = c("from", "rmvedge", "ms"), by.y = c("id", "rmvedge", "ms"), all.y = T) %>% arrange(.frame, x,y)
babystep2 <- merge(babystep1, babytweend[,-5], by.x = c("to", "ms",".frame"), by.y = c("id", "ms",".frame"), all.x=T)

idx <- is.na(babystep2$to) & babystep2$ms < 1 & babystep2$ms > 0 & babystep2$from == "V1"

babystep2$to[idx] <- "V2"
fms <- babystep2$.frame[idx]
babystep2[idx, c("x.y", "y.y")] <- babystep2[babystep2$from == "V2" & babystep2$.frame %in% fms,c("x.x", "y.x")]

ggplot(data = babystep2, aes(x = x.x, y = y.x, frame = .frame)) +
  geom_segment(aes(xend = x.y, yend = y.y, alpha = 1-ms), color = 'red') + 
  geom_point(color = 'grey40') + 
  scale_alpha_identity() + 
  facet_wrap(~.frame)
  

# bigger example 
microsteps <- ms1[1:5]
to_tween <- pretween_vertices(microsteps = microsteps, layoutparams = list(n = 50))
pte <- pretween_edges(microsteps = microsteps)

tweendat <- tween_states(data = to_tween, statelength = 18, 1, "linear",100)
tweendat$addedge <- as.logical(tweendat$addedge)
tweendat$rmvedge <- as.logical(tweendat$rmvedge)

# first, split according to microstep. 
listTweened <- split(tweendat, as.factor(tweendat$ms))

listPostTween <- list()
# then, join edges
for (i in 1:length(listTweened)){
  # if rmvedge, merge 'old' edgelist with tweened
  currentMS <- unique(listTweened[[i]]$ms)
  ms_from_edges <- floor(currentMS)
  
  if(sum(listTweened[[i]]$addedge) > 0 & currentMS != round(currentMS)) {
    step1 <- merge(microsteps[[currentMS+2]], listTweened[[i]], 
                   by.x = "from", 
                   by.y = "id", all.x = TRUE)
    names(step1)[names(step1) %in% c("x", "y")] <- c("x.from", "y.from")
    step2 <- merge(step1, listTweened[[i]], 
                   by.x = c("to", "ms", "addedge", "rmvedge",".frame"), 
                   by.y = c("id", "ms","addedge", "rmvedge",".frame"), 
                   all.x = TRUE)
    names(step2)[names(step2) %in% c("x", "y")]  <- c("x.to", "y.to")
    listPostTween[[i]] <- step2
  } else {
    step1 <- merge(microsteps[[ms_from_edges+1]], listTweened[[i]], 
                   by.x = "from", by.y = "id", all.x = TRUE)
    names(step1)[names(step1) %in% c("x", "y")] <- c("x.from", "y.from")
    step2 <- merge(step1, listTweened[[i]],
                  by.x = c("to", "ms", "addedge", "rmvedge",".frame"),
                  by.y = c("id", "ms","addedge", "rmvedge",".frame"),
                  all.x = TRUE)
    names(step2)[names(step2) %in% c("x", "y")]  <- c("x.to", "y.to")
    listPostTween[[i]] <- step2
  }
}

# merge first, then go back and overwrite. 
# use all.y so that all frames are included, and none are NA. 
#step1 <- merge(pte, tweendat, by.x = c("from","addedge", "rmvedge",  "microstep"), by.y = c("id", "addedge", "rmvedge","ms"), all.y = TRUE)
#names(step1)[names(step1) %in% c("x", "y")] <- c("x.from", "y.from")
# remove the logicals in tweened because they're only relevant to the 'from' node 
#lgcls <- names(tweendat) %in% c("addedge", "rmvedge")
#step2 <- merge(step1, tweendat[,-which(lgcls)], by.x = c("to", "microstep", ".frame"), by.y = c("id", "ms",".frame"), all.x = TRUE)
#names(step2)[names(step2) %in% c("x", "y")]  <- c("x.to", "y.to")

#B <- length(unique(pte$microstep))

# for (i in 1:B){
#   new_data_to_plot <- list()
#   msid <- unique(pte$microstep)[i]
#   nodesub <- filter(tweendat, ms == msid)
#   edgesub <- filter(pte, microstep == msid)
#   allsub <- filter(step2, microstep >= msid & microstep <= msid+1)
#   if (sum(edgesub$rmvedge) > 0){
#     fromid <- unique(edgesub$from[which(edgesub$rmvedge)])
#     toid <- unique(edgesub$to[which(edgesub$rmvedge)])
#     idx_to_change <- which(is.na(allsub$to) & allsub$microstep < msid+1 &
#                              allsub$microstep > msid & allsub$from == fromid)
#     allsub$to[idx_to_change] <- toid
#     fms <- allsub$.frame[idx_to_change]
#     allsub[idx_to_change, c("x.to", "y.to")] <- allsub[which(allsub$from == toid & 
#                                                                allsub$.frame %in% fms), c("x.from", "y.from")]
#     new_data_to_plot[[i]] <- allsub
#   } else if (sum(edges))
# }

# if (sum(listTweened[[i]]$rmvedge) > 0){
#   step1 <- merge(microsteps[[ms_from_edges+1]], listTweened[[i]], 
#                  by.x = "from", 
#                  by.y = "id", all.x = TRUE)
#   names(step1)[names(step1) %in% c("x", "y")] <- c("x.from", "y.from")
#   step2 <- merge(step1, listTweened[[i]], 
#                  by.x = c("to", "ms", "addedge", "rmvedge",".frame"), 
#                  by.y = c("id", "ms","addedge", "rmvedge",".frame"), 
#                  all.x = TRUE)
#   names(step2)[names(step2) %in% c("x", "y")]  <- c("x.to", "y.to")
#   listPostTween[[i]] <- step2
# } else if (sum(listTweened[[i]]$addedge) > 0 & ceiling(currentMS) < ms_from_edges+1){
#   
# }
# if addedge, merge 'new' edgelist with tweened

# postTweendf <- do.call("rbind", listPostTween)
# 
# testp <- ggplot(data = postTweendf, aes(x = x.from, y = y.from, xend = x.to, yend = y.to, frame = .frame, color = addedge)) +
#   geom_segment(aes(color = ms)) + 
#   geom_point(size=4, color = 'grey40') + 
#   #geom_text(aes(label=from), size = 3, alpha = .5, color = 'white') + 
#   #scale_colour_manual(guide = guides(color = "none")) + 
#   scale_alpha_identity() + 
#   theme_void() 
# animation::ani.options(interval = 1/10)
# gganimate(testp, "newmethod.gif", title_frame = F)

```


